Structured Versus Decorated Cospans

### 2020-11-13 cospans.tex
started working through chapter 3

\section{Decorated cospans}\label{DecCospansDoubleCat}
In this section we build the symmetric monoidal double category $F \lCsp$ mentioned in the introduction. Then we obtain an underlying symmetric monoidal bicategory $F \mathbf{Csp}$ using a result of Shulman \cite{Shul} and finally decategorify this to obtain a symmetric monoidal category $F\mathsf{Csp}$ which is a generalization of Fong's original decorated cospan category \cite{Fong}.  The definition of a lax monoidal pseudofunctor is recalled in \cref{sec:2cats}, and the basics of double categories are recalled in \cref{sec:doublecats}.

@@

\section{Structured versus decorated cospans} \label{EquivDoubleCats}
%\begin{comment}

In \cite{BC2}, the first two authors introduces the symmetric monoidal double category of \emph{structured cospans} as a formalism to capture open networks. One of the main goals of this paper is to provide a monoidal double equivalence between this double category and that of \emph{decorated cospans}, described in detail in \cref{DecCospansDoubleCat}. Below we recall the double category structured cospans form.

\begin{thm}\label{SC}
Given a category $\X$ with finite colimits and a category $\A$ with finite coproducts and a finite coproduct preserving functor $L \maps \A \to \X$ with $\A$ and $\X$, there exists a symmetric monoidal double category $_L \lCsp(\X)$ which has:
\begin{itemize}
\item{objects given by objects of $\A$,}
\item{vertical 1-morphisms given by morphisms of $\A$,}
\item{horizontal 1-cells $a\tickar b$ given by cospans of $\X$ of the form:
\begin{equation}\label{eq:structuredcospan}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\end{equation}}
\item 2-morphisms given by maps of cospans of $\X$ of the form:
\begin{equation}\label{eq:2cellsStrCsp}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{equation}
\end{itemize}
Composition of horizontal 1-cells and 2-morphisms is given by pushouts in $\X$ and tensoring of objects is under binary coproducts in $\A$.
\end{thm}
\begin{proof}
See the first two authors' work on structured cospans \cite{BC2}.
\end{proof}

>>>>

\section{Decorated cospans}\label{DecCospansDoubleCat}
In this section we build the symmetric monoidal double category $F \lCsp$ mentioned in the introduction. Then we obtain an underlying symmetric monoidal bicategory $F \mathbf{Csp}$ using a result of Shulman \cite{Shul} 
%!!!Christina: John suggested we might move all the horizontal bicategory and decategorified 1-category results together somewhere later/in the end.
and finally decategorify this to obtain a symmetric monoidal category $F\mathsf{Csp}$ which is a generalization of Fong's original decorated cospan category \cite{Fong}.  The definition of a lax monoidal pseudofunctor is recalled in \cref{sec:2cats}, and the basics of double categories are recalled in \cref{sec:doublecats}.

@@

\section{Structured versus decorated cospans} \label{EquivDoubleCats}

In \cite{BC2}, the first two authors introduce the symmetric monoidal double category of \emph{structured cospans} as a formalism to capture open networks. One of the main goals of this paper is to provide a monoidal double equivalence between this double category and that of \emph{decorated cospans}, described in detail in \cref{DecCospansDoubleCat}. We first recall the double category structured cospans form.

\begin{thm}\label{SC}
Given categories $\X$ and $\A$ with finite colimits and $L \maps \A \to \X$ a finitely cocontinuous functor, there exists a symmetric monoidal double category $_L\lCsp(\X)$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical morphism is a morphism of $\A$,
\item a horizontal 1-cell $a\tickar b$ is a cospan in $\X$ of the form
\begin{equation}\label{eq:structuredcospan}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B);
\end{tikzpicture}
\end{equation}
\item a 2-morphism is a map of cospans of $\X$ of the form
\begin{equation}\label{eq:2cellsStrCsp}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(a)$};
\node (B) at (1,0) {$x$};
\node (C) at (2,0) {$L(b)$};
\node (A') at (0,-1) {$L(a')$};
\node (B') at (1,-1) {$x'$};
\node (C') at (2,-1) {$L(b')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A') edge node[above]{$$} (B')
(C') edge node[above]{$$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\end{equation}
\end{itemize}
The symmetric monoidal structure is formed by using binary coproducts and the initial object.
\end{thm}

\begin{proof}
All relevant details can be found in \cite[Thm.~2.3\&3.9]{BC2}. In particular, the double category structure only requires that $\X$ have pushouts, since this is how composition is formed. Moreover, the symmetric monoidal structure requires that $\X$ also have binary coproducts and an initial object (thus all finite colimits), and that $\A$ have and $L$ preserve finite coproducts, see \cite[Thm.~3.2.3]{CourserThesis}.
\end{proof}
